---
output:
  html_document: default
  pdf_document: 
  pandoc_args: ["--lua-filter=remove-code-blocks.lua"]
---
---
fontsize: 12pt
author: "Joanna Huertas"
date: "Spring 2023"
always_allow_html: true
output:
  html_document: default
  fontsize: default
subtitle: "https://github.com/joa-kenit/EDAProject2023"
title: Evaluating the correlation between various water contaminants throughout the
  Ellerbe Creek Watershed
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

\newpage
\tableofcontents
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# Set your working directory
setwd("C:/Users/joann/Documents/EDAProject2023")

# Load your packages

library(tidyverse)
library(remotes)
library(reshape2)
library(lubridate) 
library(here) 
library(cowplot) 
library(here)
library(corrplot)

#Add pre-set themes
library(ggthemes)
library(dplyr)

#install.packages('extrafont')
library(extrafont)
font_import()
#loadfonts(device = "win")
#windowsFonts()

getwd() #Checking working directory
#here()

#install.packages("GGally")
library(GGally)

#spatial analysis
#install.packages("arcpullr")
library(arcpullr)
#install.packages("mapview")
#install.packages("maps")
library(mapview)
library(leaflet)
library(stringr)
library(sp)
library(maps)
```


# Introduction
The growing demand for freshwater coupled with human activities has resulted in increased water pollution. Water contamination is a significant environmental concern, with potential health risks to humans and wildlife. The Ellerbe Creek Watershed, located in North Carolina, is home to diverse aquatic species and provides drinking water to local communities. However, it is also vulnerable to water contamination from various sources, including agriculture, urban runoff, and industrial activities.

This project aims to evaluate the correlation between various water contaminants throughout the Ellerbe Creek Watershed. The study will investigate the presence and levels of contaminants such as heavy metals in the water, and analyze the relationship between these contaminants. The findings of this study could provide valuable insights into the impact of water contamination on the environment and human health, and help identify potential sources of pollution in the watershed.
\newpage

# Rationale and Research Questions
### Research question:
Is there a correlation between water quality parameters in the Ellerbe Creek?

### Rationale:
Understanding the correlation between different water contaminants in the Ellerbe Creek is essential to identify potential sources of pollution and implement effective management strategies to protect the environment and public health.

 A statistically significant correlation between Fe and Mn concentrations, attributable to bedrock geology, was identified.
\newpage

# Dataset Information

Our dataset was retrieved from the official Water Quality Data Web Portal of the City of Durham: http://www.durhamwaterquality.org/ 
This website contains water quality parameters of different watersheds in the City of Durham.
We have downloaded data of the Ellerbe Creek watershed, relevant for this project, and it consists of 35 water quality parameters. 

| Sampling | Ambient Sampling |
|------:|:------|
| First year of sampling    |2004|
|Frequency of Collection   |    Irregular  |
|Number of Parameters  |35|
|Types of Parameters  |Water quality measurements|
|Number of Relevant Collection Locations  |7|
|Original Goal  |Monitor safety and regulation compliance every couple of weeks|
|Data Collector  |City of Durham|

Find the metadata in /Data/Raw in the .tex file

\newpage

# Exploration Raw Data
Reading dataset:
```{r Exploratory Analysis, echo=FALSE, message=FALSE}
#Read csv
water.durham <- read.csv("./Data/Raw/ParamDurhamData_raw.csv",stringsAsFactors = TRUE)
```
Exploring dimensions and each column of the dataset:
```{r Exploratory Analysis I, echo=FALSE, message=FALSE}
#reveal dimensions, column names
colnames(water.durham)
str(water.durham)
lapply(water.durham, head)
```

\newpage

# Data Wrangling Analysis
After exploring the raw data, we have enough information and we have the need to wrangle the data to easily analyze the trends of our data, and this includes:

* Change date columns to be date objects
* Drop “NA” values
* Save the processed dataset in the Processed folder. 
* Select only the useful columns
* Adding columns for months and years

The relevant columns were selected before getting the final dataset.
This dataset was saved as “water.durham.wrangle” and used for subsequent analyses. 

These are the set of Parameters we have:
\newpage
```{r, echo=FALSE}
# extract time part from date, date column to be date object
#nchar(water.durham$Date.Time)
water.durham$Date.Time <- as.character(water.durham$Date.Time)
water.durham$Date.Time <- substr(water.durham$Date.Time,1,nchar(water.durham$Date.Time)-5)%>%
  mdy()

colnames(water.durham)[colnames(water.durham) == "Date.Time"] <- "Date"

#Select only the useful columns
water.durham.select <-  select (water.durham, id, Station.Name, Parameter, 
                                Date, Value,Unit, Sky.Condition)%>%
  drop_na(Value)

#finding unique values
unique(water.durham.select$Parameter)

#wrangle
water.durham.wrangle <- 
  water.durham.select %>%
  group_by(Date, Station.Name, Parameter, Value) %>%
  #summarise(meanAQI = mean(DAILY_AQI_VALUE)) %>%
  mutate(Month=month(Date),Year=as.factor(year(Date)))

#save the processed dataset
write.csv(water.durham.wrangle, row.names = FALSE, 
          file = "./Data/Processed/water.durham.wrangle.csv")
```

\newpage
# Exploring processed dataset
This is the summary of our dataset “water.durham.wrangle”:
```{r, echo=FALSE}
summary(water.durham.wrangle)

```

\newpage

## Question: Is there a correlation between water quality parameters in the Ellerbe Creek?
H0 : There is no significant correlation between certain parameters.
HA : There is a significant correlation between certain parameters.

The pairs of parameters chosen, based on previous studies (see references) and general knowledge, are
the following:

* Chloride and Fluoride
* BOD and DO
* Temperature and Dissolved Oxygen (DO)
* Biochemical Oxygen Demand (BOD) and Organic Carbon
* Iron and manganese
* Iron and dissolved oxygen
* Nitrate + nitrite and dissolved oxygen
* Total suspended solids + turbidity
* Total suspended solid + conductivity
* Phosphorus and nitrate
* Total Kjeldahl Nitrogen and Nitrate+Nitrite Concentration
* Chloride and total suspended solid
* Chloride and Fluoride
* Lead and pH
* pH and ammonia nitrogen

After creating a custom plot theme, we got some plots of individual parameters, a function was created to do so.

## Exploring individual trends

```{r custom plot, echo=FALSE, message=FALSE, warning=FALSE}

# Creating a custom plot theme to explore pairs of variables
my_theme <- theme_gray() + 
  theme(
    line = element_line(
      color='black',
      linewidth =1
    ),
    plot.background =   element_rect(
      color='black',
      fill='white'
    ),
    axis.text = element_text(size = 11),
    plot.title =        element_text(
      size=16,
      family="Arial",
      face = "bold",
      color='black'
    ),
    axis.title = element_text(size = 14, face = "bold"),
    legend.background = element_rect(
      color='black',
      fill = 'white'
    ),
    legend.title = element_text(size=13,  
                                family="Arial",
      color='black'
    )
  )


```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#exploring individual trends
#Alkalinity
water.alkalinity.durham <- water.durham.wrangle %>%
  filter(Parameter == "Alkalinity")%>%
    ggplot(
    mapping = aes(
      x=Date,
      y=Value,
      color=Station.Name)
    ) + 
  geom_point()+
  geom_line()+
  # Add a line of best fit and color it black
  #geom_smooth(method = "lm", se = FALSE, color = "black") +
  labs(title="Alkalinity in Water", x="Year", y="Alkalinity (mgCaCO3/L)", color="Station Name")+
  # Adjust the x and y limits to hide extreme values
  #xlim(0, 125) +
  #ylim(0, 50)+
  scale_x_date(date_breaks = "years", date_labels = "%Y")+
  my_theme
print(water.alkalinity.durham)

#nitrogen
water.nitrogen.durham <- water.durham.wrangle %>%
  filter(Parameter == "Nitrate + Nitrite as N")%>%
    ggplot(
    mapping = aes(
      x=Date,
      y=Value,
      color=Station.Name)
    ) + 
  geom_point()+
  geom_line()+
  # Add a line of best fit and color it black
  #geom_smooth(method = "lm", se = FALSE, color = "black") +
  labs(title="Nitrogen in Water", x="Year", y="Nitrogen (mg/L)", color="Station Name")+
  # Adjust the x and y limits to hide extreme values
  #xlim(0, 125) +
  #ylim(0, 50)+
  #scale_x_date(date_breaks = "years", date_labels = "%Y")+
  geom_hline(yintercept = 10, color = "red",  size = 1)+
  my_theme

print(water.nitrogen.durham)

#fluoride
water.fluoride.durham <- water.durham.wrangle %>%
  filter(Parameter == "Fluoride")%>%
    ggplot(
    mapping = aes(
      x=Date,
      y=Value,
      color=Station.Name)
    ) + 
  geom_point()+
  geom_line(color = "blue")+
  # Add a line of best fit and color it black
  #geom_smooth(method = "lm", se = FALSE, color = "black") +
  # Adjust the x and y limits to hide extreme values
  #xlim(0, 125) +
  #ylim(0, 50)+
  #scale_x_date(date_breaks = "years", date_labels = "%Y")+
  labs(title="Fluoride", x="Year", y="Fluoride (mg/L)", color="Station Name")+
  my_theme+
    geom_hline(aes(yintercept= 4, linetype = ""), colour= 'red', size = 1) +
    scale_linetype_manual(name = "MCL", values = c(1), 
                      guide = guide_legend(override.aes = list(color = c("red"))))

print(water.fluoride.durham)
```


```{r, echo=FALSE}
#creating a function for individual trends
Plot.get <- function(parameter, MCL, plottitle) {
  plot <- water.durham.wrangle %>%
  filter(Parameter == parameter)%>%
    ggplot(
    mapping = aes(
      x=Date,
      y=Value,
      color=Station.Name)
    ) + 
  geom_point()+
  geom_line(color = "blue")+
  # Add a line of best fit and color it black
  #geom_smooth(method = "lm", se = FALSE, color = "black") +
  labs(title=plottitle, x="Year", y=paste0(plottitle," ","(", pull(water.durham.wrangle[water.durham.wrangle$Parameter == "Iron" , "Unit"])[2],")"), 
       color="Station Name")+
  # Adjust the x and y limits to hide extreme values
  #xlim(0, 125) +
  #ylim(0, 50)+
  #scale_x_date(date_breaks = "years", date_labels = "%Y")+
  my_theme+
    geom_hline(aes(yintercept= MCL, linetype = ""), colour= 'red', size = 1) +
    scale_linetype_manual(name = "MCL", values = c(1), 
                      guide = guide_legend(override.aes = list(color = c("red"))))
  #Sys.sleep(1) # Scraping etiquette
  return(plot)
}

water.iron.durham <- Plot.get("Iron", 300,"Iron" )
print(water.iron.durham)
water.chromium.durham <- Plot.get("Chromium", 100,"Chromium" )
print(water.chromium.durham)
water.cad.durham <- Plot.get("Cadmium", 30,"Cadmium" )
```

## Visualization of possible correlations between 2 parameters for each available station

```{r, echo=FALSE}
#now, let's visualize possible correlations between 2 parameters
#test, this one is working with values
water.durham.wrangle %>%
    filter(Parameter == "Temperature" | Parameter == "Manganese" ) %>%
    ggplot(
      mapping = aes(
        x=Date,
        y=Value,
        color=Parameter)
    ) + 
    geom_point()+
    geom_line()+
    facet_wrap(~Station.Name, scales = "free_y") +
  scale_color_manual(values = c("red", "black")) +
  labs(title = paste0("Relationship between - ", "Temperature", " and ", "Manganese"), x = "Year", y=paste0("Value"," ","(", pull(water.durham.wrangle[water.durham.wrangle$Parameter == "Iron", "Unit"])[2],")")) +
    my_theme+
  geom_smooth(method = 'lm')

```

```{r, echo=FALSE}
#converting to a function
Plot2.get <- function(parameter1, parameter2) {
  plot <- water.durham.wrangle %>%
    filter(Parameter == parameter1 | Parameter == parameter2 ) %>%
    ggplot(
      mapping = aes(
        x=Date,
        y=Value,
        color=Parameter)) + 
    geom_point()+
    geom_line()+
    facet_wrap(~Station.Name, scales = "free_y") +
    scale_color_manual(values = c("red", "blue")) +
    labs(title = paste0("Relationship between ", parameter1, " and ", parameter2), x = "Year", y=paste0("Value"," ","(", pull(water.durham.wrangle[water.durham.wrangle$Parameter == "Iron", "Unit"])[2],")"))+
    my_theme+
    Sys.sleep(1) # Scraping etiquette
    return(plot)
}


#creating a function for a different unit system (mg/L, among others)
Plot.mg.l.get <- function(parameter1, parameter2) {
  plot <- water.durham.wrangle %>%
    filter(Parameter == parameter1 | Parameter == parameter2 ) %>%
    ggplot(
      mapping = aes(
        x=Date,
        y=Value,
        color=Parameter)) + 
    geom_point()+
    geom_line()+
    facet_wrap(~Station.Name, scales = "free_y") +
    scale_color_manual(values = c("red", "blue")) +
    labs(title = paste0("Relationship between ", parameter1, " and ", parameter2), x = "Year", y=paste0("Value"," ","(", pull(water.durham.wrangle[water.durham.wrangle$Parameter == "Chloride", "Unit"])[2],")"))+
    my_theme+
    Sys.sleep(1) # Scraping etiquette
    return(plot)
}

#creating a function for a different unit system (mg/L and C°)
Plot.mg.l.c.get <- function(parameter1, parameter2) {
  plot <- water.durham.wrangle %>%
    filter(Parameter == parameter1 | Parameter == parameter2 ) %>%
    ggplot(
      mapping = aes(
        x=Date,
        y=Value,
        color=Parameter)) + 
    geom_point()+
    geom_line()+
    facet_wrap(~Station.Name, scales = "free_y") +
    scale_color_manual(values = c("red", "blue")) +
    labs(title = paste0("Relationship between ", parameter1, " and ", parameter2), x = "Year", y=paste0("Value (C°)"," ","(", pull(water.durham.wrangle[water.durham.wrangle$Parameter == "Chloride", "Unit"])[2],")"))+
    my_theme+
    Sys.sleep(1) # Scraping etiquette
    return(plot)
}
```
\newpage
```{r, echo=FALSE}
#relationship Chloride and Fluoride
water.cl.f <- Plot.mg.l.get("Chloride", "Fluoride")
print(water.cl.f)

#relationship BOD and DO
water.bod.do <- Plot.mg.l.get("Biochemical Oxygen Demand", "Dissolved Oxygen")
print(water.bod.do)

#relationship Temperature and Dissolved Oxygen (DO)
water.temp.do <- Plot.mg.l.c.get("Temperature", "Dissolved Oxygen")
print(water.temp.do)

#relationship Biochemical Oxygen Demand (BOD) and Organic Carbon
water.bod.oc <- Plot.mg.l.get("Biochemical Oxygen Demand", "Organic Carbon")
print(water.bod.oc)

#relationship iron and manganese
water.iron.mn <- Plot2.get("Iron", "Manganese")
print(water.iron.mn)

#relationship iron and dissolved oxygen
water.iron.do <- Plot2.get("Iron", "Dissolved Oxygen")
print(water.iron.do)

#relationship nitrate + nitrite and dissolved oxygen
water.nit.do <- Plot.mg.l.get("Nitrate + Nitrite as N", "Dissolved Oxygen")
print(water.nit.do)

#relationship total suspended solids + turbidity
water.tss.turb <- Plot.mg.l.get("Total Suspended Solids", "Turbidity")
print(water.tss.turb)

#relationship total suspended solid + conductivity
water.tss.cond <- Plot.mg.l.get("Total Suspended Solids", "Conductivity")
print(water.tss.cond)

#relationship phosphorus and nitrate
water.p.nit <- Plot.mg.l.get("Total Phosphorus", "Nitrate + Nitrite as N")
print(water.p.nit)

#relationship Total Kjeldahl Nitrogen and Nitrate+Nitrite Concentration
water.totnit.nit <- Plot.mg.l.get("Total Kjeldahl Nitrogen", "Nitrate + Nitrite as N")
print(water.totnit.nit)

#relationship Chloride and total suspended solid
water.cl.tss <- Plot.mg.l.get("Chloride", "Total Suspended Solids")
print(water.cl.tss)

#relationship lead and pH
water.pb.ph <- Plot2.get("Lead", "pH")
print(water.pb.ph)

#relationship pH and ammonia nitrogen
water.ph.n <- Plot2.get("pH", "Ammonia Nitrogen")
print(water.ph.n)
```
\newpage

## Visualization
Visualizing the relationship among two variables with a scatter plot of chloride by
fluoride.

```{r echo=FALSE, warning=FALSE}

#long to wide format
wide_data <- dcast(water.durham.wrangle, Year + Month + Station.Name ~ Parameter, fun.aggregate = mean,
                   value.var= "Value")

water_wide_plot <- 
  ggplot(wide_data, aes(x=Chloride, y=Fluoride, color = Station.Name))+
  geom_point()+
  geom_smooth(method="lm", se = FALSE, color = "blue") +
  #xlim(0, 125) +
  #ylim(0, 35)+
  labs(title= "Chloride vs Fluoride", x= "Chloride (mg/L)", y= "Fluoride (mg/L)", color="Station.Name")+
   theme(legend.text = element_text(size = 7), legend.position = "top", 
         legend.justification = "right", legend.key.width = unit(10, "pt"),
         legend.key.height = unit(1, "cm"))

print(water_wide_plot)
```
In this case, the scatterplot suggests that there could be a negative relationship between chloride and fluoride.
\newpage
## Linear Regression Analysis
```{r, echo=FALSE, warning=FALSE}
#changing name of the column for further analysis
colnames(wide_data) <-  make.names(names(wide_data))

#regression analysis
cl.f.regression <- lm(data = wide_data, Chloride ~ Fluoride)
summary(cl.f.regression)

bod.do.regression <- lm(data = wide_data, Biochemical.Oxygen.Demand ~ Dissolved.Oxygen)
summary(bod.do.regression)

temp.do.regression <- lm(data = wide_data, Temperature  ~ Dissolved.Oxygen)
summary(temp.do.regression)

bod.oc.regression <- lm(data = wide_data, Biochemical.Oxygen.Demand ~ Organic.Carbon)
summary(bod.oc.regression)

fe.mn.regression <- lm(data = wide_data, Iron ~ Dissolved.Oxygen)
summary(fe.mn.regression)

fe.f.regression <- lm(data = wide_data, Iron ~ Fluoride)
summary(fe.f.regression)

n.do.regression <- lm(data = wide_data, Nitrate...Nitrite.as.N ~ Dissolved.Oxygen)
summary(n.do.regression)

tss.turbidity.regression <- lm(data = wide_data, Total.Suspended.Solids ~ Turbidity)
summary(tss.turbidity.regression)

tss.conductivity.regression <- lm(data = wide_data, Total.Suspended.Solids ~ Conductivity)
summary(tss.conductivity.regression)

tp.n.regression <- lm(data = wide_data, Total.Phosphorus ~ Nitrate...Nitrite.as.N)
summary(tp.n.regression)

tkn.n.regression <- lm(data = wide_data, Total.Kjeldahl.Nitrogen ~ Nitrate...Nitrite.as.N)
summary(tkn.n.regression)

cl.tss.regression <- lm(data = wide_data, Chloride ~ Total.Suspended.Solids)
summary(cl.tss.regression)

pb.ph.regression <- lm(data = wide_data, Lead ~ pH)
summary(pb.ph.regression)

ph.ammonia.regression <- lm(data = wide_data, pH ~ Ammonia.Nitrogen)
summary(ph.ammonia.regression)

```
\newpage

## Interpretation of results

```{r, echo=FALSE}
df.residual(summary(cl.f.regression))

r_squared <- paste("The linear regression analysis indicates that changes in " ,"parameter1 ", "explain ", round(100*summary(cl.f.regression)$r.squared, 2), 
"%", " of the variability in ", "parameter2. ", "This finding is based on:", " degrees of freedom = ", summary(cl.f.regression)$df[1], ", ", 
summary(cl.f.regression)$df[2], "; degrees of freedom for the residuals = ", summary(cl.f.regression)$df.residual,
"; standard error = ", round(summary(cl.f.regression)$sigma, 2), "; p-value = ", formatC(summary(cl.f.regression)$coefficients[2, 4], digits = 2, format = "f"), sep = "")

#print(r_squared)

```
\newpage
```{r, echo=FALSE}
#creating a function based on previous test

# Define a function to format the regression results
format_regression_results <- function(parameter1, parameter2, regression) {
  
  # Extract the regression results
  summary <- summary(regression)
  r_squared <- round(100 * summary$r.squared, 2)
  df1 <- summary$df[1]
  df2 <- summary$df[2]
  se <- round(summary$sigma, 2)
  p_value <- formatC(summary$coefficients[2, 4], digits = 2, format = "f")
  
  # Build the formatted string
  result <- paste0(
    "The linear regression analysis indicates that changes in ", parameter1, 
    " explain ", r_squared, "% of the variability in ", parameter2, 
    ". This finding is based on ", df1, " degrees of freedom for the numerator, ", 
    df2, " degrees of freedom for the denominator, a standard error of ", se, 
    ", and a p-value of ", p_value, "."
  )
  
  # Add the significance information
  if (summary$coefficients[2, 4] < 0.05) {
    result <- paste(result, " (statistically significant)", sep = "")
  } else {
    result <- paste(result, " (not statistically significant)", sep = "")
  }
  
  return(result)
}

# linear regression analysis findings
fit <- lm( Chloride ~ Fluoride, data = wide_data)
result <- format_regression_results("Chloride", "Fluoride", cl.f.regression)
print(result)


fit <- lm( Biochemical.Oxygen.Demand ~ Dissolved.Oxygen, data = wide_data)
result <- format_regression_results("Biochemical.Oxygen.Demand", "Dissolved.Oxygen", bod.do.regression)
print(result)

fit <- lm( Temperature ~ Dissolved.Oxygen, data = wide_data)
result <- format_regression_results("Temperature", "Dissolved.Oxygen", temp.do.regression)
print(result)

fit <- lm( Biochemical.Oxygen.Demand ~ Organic.Carbon, data = wide_data)
result <- format_regression_results("Biochemical.Oxygen.Demand", "Organic.Carbon", bod.oc.regression)
print(result)

fit <- lm( Iron ~ Dissolved.Oxygen, data = wide_data)
result <- format_regression_results("Iron", "Dissolved.Oxygen", fe.mn.regression)
print(result)

fit <- lm( Iron ~ Fluoride, data = wide_data)
result <- format_regression_results("Iron", "Fluoride", fe.f.regression)
print(result)

fit <- lm( Nitrate...Nitrite.as.N ~ Dissolved.Oxygen, data = wide_data)
result <- format_regression_results("Nitrate...Nitrite.as.N", "Dissolved.Oxygen", n.do.regression)
print(result)

fit <- lm( Total.Suspended.Solids ~ Turbidity, data = wide_data)
result <- format_regression_results("Total.Suspended.Solids", "Turbidity", tss.turbidity.regression)
print(result)

fit <- lm( Total.Suspended.Solids ~ Conductivity, data = wide_data)
result <- format_regression_results("Total.Suspended.Solids", "Conductivity", tss.conductivity.regression)
print(result)

fit <- lm( Total.Phosphorus ~ Nitrate...Nitrite.as.N, data = wide_data)
result <- format_regression_results("Total.Phosphorus", "Nitrate...Nitrite.as.N", tp.n.regression)
print(result)

fit <- lm( Total.Kjeldahl.Nitrogen ~ Nitrate...Nitrite.as.N, data = wide_data)
result <- format_regression_results("Total.Kjeldahl.Nitrogen", "Nitrate...Nitrite.as.N", tkn.n.regression)
print(result)

fit <- lm( Chloride ~ Total.Suspended.Solids, data = wide_data)
result <- format_regression_results("Chloride", "Total.Suspended.Solids", cl.tss.regression)
print(result)

fit <- lm( Lead ~ pH, data = wide_data)
result <- format_regression_results("Lead", "pH", pb.ph.regression)
print(result)

fit <- lm( Ammonia.Nitrogen ~ pH, data = wide_data)
result <- format_regression_results("Ammonia.Nitrogen", "pH", ph.ammonia.regression)
print(result)
```

\newpage
# Spatial Analysis
This is the location of the main stations along the Ellerbe Creek Watershed:
```{r, echo=FALSE, warning=FALSE}
#reading stations dataset
stationData = read.csv('./Data/Raw/durham_station.csv')
stationData1 = stationData[stationData$Name %in% c("EL5.5GC","EL1.9EC","EL5.6EC","EL8.1GC","EL8.5SEC","EL7.1SEC","EL8.6SECUT","EL7.1EC"),]

#getting spatial layer
huc14 <- get_spatial_layer(url = "https://services1.arcgis.com/XBhYkoXKJCRHbe7M/arcgis/rest/services/Ellerbe_Creek_CatchmentsWMIP_view/FeatureServer/0") 
huc15 <- mapview(huc14)

#using leaflet to draw the map
leaflet() %>%
  addTiles('https://server.arcgisonline.com/ArcGIS/rest/services/Canvas/World_Light_Gray_Base/MapServer/tile/{z}/{y}/{x}') %>%
  addPolygons(data = huc14, weight = 1, opacity = 1, color = "black", fillOpacity = 0)%>% addProviderTiles(providers$Esri.NatGeoWorldMap)%>%addMarkers(data = stationData1, lng = ~Longitude, lat = ~Latitude, label = ~Name)
```


\newpage

# Summary and Conclusions
* For most of the analyses, the percentage of variability explained by the independent variable was found to be less than 5%. 
* However, for some analyses such as the relationship between Temperature and Dissolved Oxygen or Total Suspended Solids and Turbidity, the percentage of variability explained was found to be higher, at 57.81% and 68.72%, respectively. In 7 out of 14 cases, the p-value was found to be statistically significant, indicating a relationship between the two variables, while in others it was not. 


\newpage

# References
* City of Durham, water quality data—Web portal. (n.d.). Retrieved 1 May 2023, from http://www.durhamwaterquality.org/
* Chigor, V. N., Umoh, V. J., Okuofu, C. A., Ameh, J. B., Igbinosa, E. O., & Okoh, A. I. (2012). Water quality assessment: Surface water sources used for drinking and irrigation in Zaria, Nigeria are a public health hazard. Environmental Monitoring and Assessment, 184(5), 3389–3400. https://doi.org/10.1007/s10661-011-2396-9
* O’Donnell, D. (2021, September 20). Three main types of water quality parameters explained. Sensorex. https://sensorex.com/three-main-types-of-water-quality-parameters-explained/
* Zaw, M., & Chiswell, B. (1999). Iron and manganese dynamics in lake water. Water Research, 33(8), 1900–1910. https://doi.org/10.1016/S0043-1354(98)00360-1
* Johnson, C. D., Nandi, A., Joyner, T. A., & Luffman, I. (2018). Iron and manganese in groundwater: Using kriging and gis to locate high concentrations in buncombe county, north carolina. Groundwater, 56(1), 87–95. https://doi.org/10.1111/gwat.12560

